[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks

### 2018-03-13 article/Caze_2018.tex
Plot raster plot with different dt's for high connection strengths, various changes to text

Our replication does not reproduce all the quantitative details of the original plots. The parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article. For linear networks, we do observe almost no simulations classified in this category, while the original article has a thin line of such classifications between the stable regime without propagation (green) and the regime where simulations are unstable even without stimulation (red). Similarly, in the original article the parameter exploration shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} where the network is unstable only after the synchronous stimulation (yellow). In contrast, our simulations show a much smaller zone when the inhibitory connection strength is small (Fig.~\ref{fig:grid}B). It seems that these differences mostly stem from the fact that the transition to the regime where simulations are unstable without stimulation (red) happens for lower excitatory and higher inhibitory connection strength in our simulations than in the original article, i.e. that the red zone is shifted towards the bottom left. This is particularly evident for strong excitatory and inhibitory connections in non-linearly coupled networks (bottom right in Fig.~\ref{fig:grid}B and C) which are unstable in our simulations, but stable or only unstable after stimulation in the original publication. Finally, non-linearly coupled networks show a discontinuity for weak inhibitory connection strengths in the original publication. For increasing excitatory connection strengths, network behaviour goes from stable (blue) to unstable after stimulation (yellow) to unstable before stimulation (red), but then get classified again as stable (blue) in a small range of parameters before finally transitioning to unstable activity (yellow and red). We do not observe this in our simulations.

Despite these differences, the qualitative network behaviour looks very similar on a coarse scale. Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not. The differences in our replication are most likely due to our use of a clock-based simulation method  (see Fig.~\ref{fig:grid_detail}). In principle, another reason could be implementation details of the initial ``spikes in transit'' that were not fully specified in the original article, but this seems unlikely as completely removing them did not change the results in a noticeable way (not shown).

@@

We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences. The grid search differs in certain regions and the markov analysis demonstrate a slight shifth upward for the numerical results. These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision. We tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained qualitatively the same results (not shown). Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.

@@

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov_correction}. Without correction (see text)}
\end{figure}
\end{document}

>>>>

Our replication does not reproduce all the quantitative details of the original plots. The parameter range where simulations are only unstable after they receive the synchronous stimulation (yellow) is smaller than in the original article. For linear networks, we do observe almost no simulations classified in this category, while the original article has a thin line of such classifications between the stable regime without propagation (green) and the regime where simulations are unstable even without stimulation (red). Similarly, in the original article the parameter exploration shows for each inhibitory connection weight a range of excitatory connection weights of about \SI{5}{\milli\volt} where the network is unstable only after the synchronous stimulation (yellow). In contrast, our simulations show a much smaller zone when the inhibitory connection strength is small (Fig.~\ref{fig:grid}B). It seems that these differences mostly stem from the fact that the transition to the regime where simulations are unstable without stimulation (red) happens for lower excitatory and higher inhibitory connection strength in our simulations than in the original article, i.e. that the red zone is shifted towards the bottom left. This is particularly evident for strong excitatory and inhibitory connections in non-linearly coupled networks (bottom right in Fig.~\ref{fig:grid}B and C) which are unstable in our simulations, but stable or only unstable after stimulation in the original publication. Simulating the parameter combination corresponding to the point in the bottom right corner of the plot different time resolutions (Fig.~\ref{fig:raster_strong}), we find that this difference does indeed seem to be triggered by limited time resolution in the model. The forced synchronisation of all spikes to a grid of \SI{0.1}{\milli\second}, together with the homogeneity of the synaptic delays, leads to strong oscillations in our simulation (Fig.~\ref{fig:raster_strong}A), which is consequently classified as unstable. When we use a simulation time step of \SI{0.01}{\milli\second}, these osciallations disappear and the network activity therefore gets classified as ``stable without persistent propagation''.  Finally, non-linearly coupled networks show a discontinuity for weak inhibitory connection strengths in the original publication. For increasing excitatory connection strengths, network behaviour goes from stable (blue) to unstable after stimulation (yellow) to unstable before stimulation (red), but then get classified again as stable (blue) in a small range of parameters before finally transitioning to unstable activity (yellow and red). We do not observe this in our simulations, but simulations with a finer time resolution (Fig.~\ref{fig:grid_detail}C) indicate that this difference could be caused by the limited time resolution of our simulations as well.

Despite these differences, the qualitative network behaviour looks very similar on a coarse scale. Importantly, it strongly supports the original paper's main conclusion that non-linearly coupled networks have a large parameter range of stable propagation (i.e., blue-coloured points), while linearly coupled networks do not. The differences in our replication are most likely due to our use of a clock-based simulation method with a limited time resolution (see Fig.~\ref{fig:grid_detail} and Fig.~\ref{fig:raster_strong}). In principle, another reason could be implementation details of the initial ``spikes in transit'' that were not fully specified in the original article, but this seems unlikely since completely removing them did not change the results in a noticeable way (not shown).

@@

We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences. The grid search over parameters showed a higher prevalence of networks with unstable activity before the onset of stimulation compared to the original study. In the analysis of the evolution of synchronous pulses of different sizes, our results need a correction for the use of a clock-based simulation method to show a good quantative match with the original results. However, a marked difference remains for small initial group sizes. The differences we see are most likely due to our use of a clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision. However, the main qualitative finding---non-linearly coupled networks have a large parameter range supporting persistent propagation of synchronous activity whereas linearly coupled networks do not---remains if we use finer (\SI{0.05}{\milli\second}) or coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions (data not shown).

@@

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_10us.pdf}
\end{subfigure}
\caption{\label{fig:raster_strong}
\textbf{Propagation of synchronous spiking in non-linearly coupled networks with strong recurrent connections when simulated with a time step of \SI{0.1}{\milli\second} (A), or \SI{0.01}{\milli\second} (B).} All conventions as in Fig.~\ref{fig:raster}, but for strong recurrent connections ($w_\text{ex} = w_\text{in} = \SI{0.4}{\milli\volt}$). Both simulated networks use non-linearly coupling, the network shown in panel (B) is simulated with a 10-times finer temporal resolution compared to the other results shown.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov_corrected}. The numerical results shown here (gray lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
\end{figure}
\end{document}
